---
title: "Gov 50 Final Project"
author: "James McAffrey"
description: "Education Data"
output:
  distill::distill_article:
    self_contained: false
---


## Project Data Proposal

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)


covid <- read.csv("data/covid-19-polls-master/covid_approval_polls.csv")

view(covid)
```


```{r}
covid |> 
  count(subject)
```
\newpage
## Step 1
Creates a dataset avg_covid that gets the average of the approval and disaproval ratings for both Trump and Biden.
```{r}
avg_covid <- covid |>
  group_by(party, subject) |>
  summarize(
    avg_approve = mean(approve, na.rm = TRUE),
    avg_disapprove = mean(disapprove, na.rm = TRUE))
```
##Step 2
Filters the data to only Trump's information. Then creates two graphs: avg_t_support and avg_t_disapprove. Each showing the average percent of the population who approves and disapproves for Trump sorted by party and total.
```{r}
avg_t <- avg_covid |>
  filter(subject == "Trump")

avg_t_support <- ggplot(data = avg_t, mapping = aes(x = party, y = avg_approve)) +
  geom_col() +
  labs(
    title = "Trump Approval",
    y = "Percent of Population who Approves",
    x = "Party"
  )

avg_t_disapprove <- ggplot(data = avg_t, mapping = aes(x = party, y = avg_disapprove)) +
  geom_col() +
  labs(
    title = "Trump Disapproval",
    y = "Percent of Population who Approves",
    x = "Party"
  )

```
\newpage
```{r}
avg_t_support
avg_t_disapprove

```
##Step 3
Filters the data to only Biden's information. Then creates two graphs: avg_b_support and avg_b_disapprove. Each showing the average percent of the population who approves and disapproves for Biden sorted by party and total.
```{r}
avg_b <- avg_covid |>
  filter(subject == "Biden")

avg_b_support <- ggplot(data = avg_b, mapping = aes(x = party, y = avg_approve)) +
  geom_col() +
   labs(
    title = "Biden Approval",
    y = "Percent of Population who Approves",
    x = "Party"
  )

avg_b_disapprove <- ggplot(data = avg_b, mapping = aes(x = party, y = avg_disapprove)) +
  geom_col() +
  labs(
    title = "Biden Disapproval",
    y = "Percent of Population who Disapproves",
    x = "Party"
  )
```
\newpage
```{r}
avg_b_support
avg_b_disapprove
```














